---
title: "NCC ESCC_Natera analysis 05232024"
output: html_notebook
---
library(swimplot)
library(grid)
library(gtable)
library(readr) 
library(mosaic)
library(dplyr) 
library(survival) 
library(survminer) 
library(ggplot2)
library(scales)
library(coxphf)
library(ggthemes)
library(tidyverse)
library(gtsummary)
library(flextable)
library(parameters)
library(car)
library(ComplexHeatmap)
library(tidyverse)
library(readxl)
library(survival)
library(janitor)
library(DT)

#ctDNA Detection Rates by Window and Stages
```{r}
#ctDNA at Baseline
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data$ctDNA.Baseline <- factor(circ_data$ctDNA.Baseline, levels=c("NEGATIVE","POSITIVE"))
circ_data <- subset(circ_data, ctDNA.Baseline %in% c("NEGATIVE", "POSITIVE"))
circ_data$Stage <- factor(circ_data$Stage, levels=c("I","II", "III","IV"))
positive_counts_by_stage <- aggregate(circ_data$ctDNA.Baseline == "POSITIVE", by=list(circ_data$Stage), FUN=sum)
total_counts_by_stage <- aggregate(circ_data$ctDNA.Baseline, by=list(circ_data$Stage), FUN=length)
combined_data <- data.frame(
  Stage = total_counts_by_stage$Group.1,
  Total_Count = total_counts_by_stage$x,
  Positive_Count = positive_counts_by_stage$x,
  Rate = (positive_counts_by_stage$x / total_counts_by_stage$x) * 100  # Convert to percentage
)
combined_data$Rate <- sprintf("%.2f%%", combined_data$Rate)
overall_total_count <- nrow(circ_data)
overall_positive_count <- nrow(circ_data[circ_data$ctDNA.Baseline == "POSITIVE",])
overall_positivity_rate <- (overall_positive_count / overall_total_count) * 100  # Convert to percentage
overall_row <- data.frame(
  Stage = "Overall",
  Total_Count = overall_total_count,
  Positive_Count = overall_positive_count,
  Rate = sprintf("%.2f%%", overall_positivity_rate)
)
combined_data <- rbind(combined_data, overall_row)
print(combined_data)

#ctDNA post-NAC
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data$ctDNA.postnac <- factor(circ_data$ctDNA.postnac, levels=c("NEGATIVE","POSITIVE"))
circ_data <- subset(circ_data, ctDNA.postnac %in% c("NEGATIVE", "POSITIVE"))
circ_data$Stage <- factor(circ_data$Stage, levels=c("I","II", "III","IV"))
positive_counts_by_stage <- aggregate(circ_data$ctDNA.postnac == "POSITIVE", by=list(circ_data$Stage), FUN=sum)
total_counts_by_stage <- aggregate(circ_data$ctDNA.postnac, by=list(circ_data$Stage), FUN=length)
combined_data <- data.frame(
  Stage = total_counts_by_stage$Group.1,
  Total_Count = total_counts_by_stage$x,
  Positive_Count = positive_counts_by_stage$x,
  Rate = (positive_counts_by_stage$x / total_counts_by_stage$x) * 100  # Convert to percentage
)
combined_data$Rate <- sprintf("%.2f%%", combined_data$Rate)
overall_total_count <- nrow(circ_data)
overall_positive_count <- nrow(circ_data[circ_data$ctDNA.postnac == "POSITIVE",])
overall_positivity_rate <- (overall_positive_count / overall_total_count) * 100  # Convert to percentage
overall_row <- data.frame(
  Stage = "Overall",
  Total_Count = overall_total_count,
  Positive_Count = overall_positive_count,
  Rate = sprintf("%.2f%%", overall_positivity_rate)
)
combined_data <- rbind(combined_data, overall_row)
print(combined_data)

#ctDNA at MRD Window
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.MRD!="",]
circ_data$ctDNA.MRD <- factor(circ_data$ctDNA.MRD, levels=c("NEGATIVE","POSITIVE"))
circ_data$Stage <- factor(circ_data$Stage, levels=c("I","II", "III","IV"))
positive_counts_by_stage <- aggregate(circ_data$ctDNA.MRD == "POSITIVE", by=list(circ_data$Stage), FUN=sum)
total_counts_by_stage <- aggregate(circ_data$ctDNA.MRD, by=list(circ_data$Stage), FUN=length)
combined_data <- data.frame(
  Stage = total_counts_by_stage$Group.1,
  Total_Count = total_counts_by_stage$x,
  Positive_Count = positive_counts_by_stage$x,
  Rate = (positive_counts_by_stage$x / total_counts_by_stage$x) * 100  # Convert to percentage
)
combined_data$Rate <- sprintf("%.2f%%", combined_data$Rate)
overall_total_count <- nrow(circ_data)
overall_positive_count <- nrow(circ_data[circ_data$ctDNA.MRD == "POSITIVE",])
overall_positivity_rate <- (overall_positive_count / overall_total_count) * 100  # Convert to percentage
overall_row <- data.frame(
  Stage = "Overall",
  Total_Count = overall_total_count,
  Positive_Count = overall_positive_count,
  Rate = sprintf("%.2f%%", overall_positivity_rate)
)
combined_data <- rbind(combined_data, overall_row)
print(combined_data)

#ctDNA at Surveillance Window
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data$ctDNA.Surveillance <- factor(circ_data$ctDNA.Surveillance, levels=c("NEGATIVE","POSITIVE"))
circ_data <- subset(circ_data, ctDNA.Surveillance %in% c("NEGATIVE", "POSITIVE"))
circ_data$Stage <- factor(circ_data$Stage, levels=c("I","II", "III","IV"))
positive_counts_by_stage <- aggregate(circ_data$ctDNA.Surveillance == "POSITIVE", by=list(circ_data$Stage), FUN=sum)
total_counts_by_stage <- aggregate(circ_data$ctDNA.Surveillance, by=list(circ_data$Stage), FUN=length)
combined_data <- data.frame(
  Stage = total_counts_by_stage$Group.1,
  Total_Count = total_counts_by_stage$x,
  Positive_Count = positive_counts_by_stage$x,
  Rate = (positive_counts_by_stage$x / total_counts_by_stage$x) * 100  # Convert to percentage
)
combined_data$Rate <- sprintf("%.2f%%", combined_data$Rate)
overall_total_count <- nrow(circ_data)
overall_positive_count <- nrow(circ_data[circ_data$ctDNA.Surveillance == "POSITIVE",])
overall_positivity_rate <- (overall_positive_count / overall_total_count) * 100  # Convert to percentage
overall_row <- data.frame(
  Stage = "Overall",
  Total_Count = overall_total_count,
  Positive_Count = overall_positive_count,
  Rate = sprintf("%.2f%%", overall_positivity_rate)
)
combined_data <- rbind(combined_data, overall_row)
print(combined_data)
```



#Summary Table
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_datadf <- as.data.frame(circ_data)

circ_data_subset <- circ_data %>%
  select(
    Gender,
    Age,
    PS,
    Stage,
    TNM.Detailed,
    Neo.Regiment,
    ypStage,
    ypTNM.Detailed,
    TRG,
    pCR,
    Adj.Regimen,
    RFS.Event,
    OS.Event,
    RFS.months,
    OS.months) %>%
  mutate(
    Gender = factor(Gender),
    Age = as.numeric(Age),
    PS = factor(PS),
    Stage = factor(Stage),
    TNM.Detailed = factor(TNM.Detailed),
    Neo.Regiment = factor(Neo.Regiment),
    ypStage = factor(ypStage),
    ypTNM.Detailed = factor(ypTNM.Detailed),
    TRG = factor(TRG),
    pCR = factor(pCR),
    Adj.Regimen = factor(Adj.Regimen),
    RFS.Event = factor(RFS.Event, levels = c("FALSE", "TRUE"), labels = c("No Recurrence", "Recurrence")),
    OS.Event = factor(OS.Event, levels = c("FALSE", "TRUE"), labels = c("Alive", "Deceased")),
    RFS.months = as.numeric(RFS.months),
    OS.months = as.numeric(OS.months))
table1 <- circ_data_subset %>%
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{median} ({min} - {max})",
      all_categorical() ~ "{n} ({p}%)")) %>%
  bold_labels()
table1
fit1 <- as_flex_table(
  table1,
  include = everything(),
  return_calls = FALSE,
  strip_md_bold = TRUE)
fit1
save_as_docx(fit1, path= "~/Downloads/table1.docx")
```



#Heatmap for the clinical factors
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_datadf <- as.data.frame(circ_data)
circ_data <- circ_data %>% arrange(Stage)
circ_datadf <- as.data.frame(circ_data)

ha <- HeatmapAnnotation(
  Stage = circ_data$Stage,
  Gender = circ_data$Gender,
  Neo.Regiment = circ_data$Neo.Regiment,
  TRG = circ_data$TRG,
  pCR = circ_data$pCR,
  ypStage = circ_data$ypStage,
  Adj.Regimen = circ_data$Adj.Regimen,
  ctDNA.Baseline = circ_data$ctDNA.Baseline,
  ctDNA.postnac = circ_data$ctDNA.postnac,
  ctDNA.4w = circ_data$ctDNA.4w,
  ctDNA.12w = circ_data$ctDNA.12w,
  ctDNA.24w = circ_data$ctDNA.24w,
  ctDNA.36w = circ_data$ctDNA.36w,
  ctDNA.48w = circ_data$ctDNA.48w,
  ctDNA.Surveillance = circ_data$ctDNA.Surveillance,
  RFS.Event = circ_data$RFS.Event,
  OS.Event = circ_data$OS.Event,
  
    col = list(Stage = c("I" = "seagreen1", "II" = "khaki", "III" = "orange", "IV" = "darkmagenta"),
    Gender = c("FEMALE" = "goldenrod" , "MALE" = "blue4"),
    Neo.Regiment = c("FOLFOX6" = "coral", "FP" ="darkgreen", "DCF" ="brown"),
    TRG = c("1" = "yellow3", "2" ="darkgreen", "3" = "brown2"),
    pCR = c("TRUE" = "lightblue", "FALSE" ="orange"),
    ypStage = c("0" = "khaki","I" = "seagreen2", "II" = "cornflowerblue","III" = "orange","IV" ="darkmagenta"),
    Adj.Regimen = c("Nivolumab" = "darkmagenta"),
    ctDNA.Baseline = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.postnac = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.4w = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.12w = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.24w = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.36w = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.48w = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    ctDNA.Surveillance = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
    RFS.Event = c("TRUE" = "red3", "FALSE" ="blue"),
    OS.Event = c("TRUE" = "black", "FALSE" ="grey")
)
)
ht <- Heatmap(matrix(nrow = 0, ncol = length(circ_data$cStage)),show_row_names = FALSE,cluster_rows = F,cluster_columns = FALSE, top_annotation = ha)
pdf("heatmap.pdf",width = 7, height = 7)
draw(ht, annotation_legend_side = "bottom")
dev.off()
```

#Overview plot
```{r}
setwd("~/Downloads")
clinstage<- read.csv("NCC_ESCC_OP.csv")
clinstage_df<- as.data.frame(clinstage)

#Display the swimmer plot with the label box
oplot<-swimmer_plot(df=clinstage_df,
                    id='PatientName',
                    end='fu.diff.months',
                    fill='gray',
                    width=.01,)
oplot <- oplot + theme(panel.border = element_blank())
oplot <- oplot + scale_y_continuous(breaks = seq(-12, 108, by = 6))
oplot <- oplot + labs(x ="Patients" , y="Months from Surgery")
oplot

##plot events
oplot_ev1 <- oplot + swimmer_points(df_points=clinstage_df,
                                    id='PatientName',
                                    time='date.diff.months',
                                    name_shape ='Event_type',
                                    name_col = 'Event',
                                    size=3.5,fill='black',
                                    #col='darkgreen'
)
oplot_ev1

#Shape customization to Event_type

oplot_ev1.1 <- oplot_ev1 + ggplot2::scale_shape_manual(name="Event_type",values=c(1,16,6,18,18,4),breaks=c('ctDNA_neg','ctDNA_pos', 'Imaging','Surgery','Biopsy', 'Death'))

oplot_ev1.1

#plot treatment

oplot_ev2 <- oplot_ev1.1 + swimmer_lines(df_lines=clinstage_df,
                                         id='PatientName',
                                         start='Tx_start.months',
                                         end='Tx_end.months',
                                         name_col='Tx_type',
                                         size=3.5,
                                         name_alpha = 1.0)
oplot_ev2 <- oplot_ev2 + guides(linetype = guide_legend(override.aes = list(size = 5, color = "black")))
oplot_ev2  


#colour customization
#orange=NAC chemo,orangered=NAC ChemoIO,  purple=ACT, Black=Death, Red=PD, ctDNA negative=white, ctDNA positive=black, cystectomy=blue, TURBT=gray, ACT Chemo=purple, ACT IO=lightblue 
oplot_ev2.2 <- oplot_ev2 + ggplot2::scale_color_manual(name="Event",values=c("lightblue", "gray", "black", "black", "orange", "green", "red", "blue"))
oplot_ev2.2
```

#RFS by ctDNA post-NAT
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.postnac!="",]
circ_datadf <- as.data.frame(circ_data)

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.postnac, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.postnac, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="RFS - ctDNA post-NAC", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", legend.labs=c("ctDNA Negative", "ctDNA Positive"), legend.title="")
summary(KM_curve, times= c(12, 24))
circ_data$ctDNA.postnac <- factor(circ_data$ctDNA.postnac, levels=c("NEGATIVE","POSITIVE"))
cox_fit <- coxph(surv_object ~ ctDNA.postnac, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)
```


#RFS by ctDNA Clearance post-NAC - 3 Groups
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.Clearance!="",]
circ_datadf <- as.data.frame(circ_data)

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.Clearance, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.Clearance, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("red","blue","green"), title="RFS - ctDNA Clearance post-NAC", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", legend.labs=c("No Clearance", "Sustained", "Transient"), legend.title="")
summary(KM_curve, times= c(12, 18, 24))
circ_data$ctDNA.Clearance <- factor(circ_data$ctDNA.Clearance, levels=c("Sustained","Transient", "No Clearance"))
cox_fit <- coxph(surv_object ~ ctDNA.Clearance, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
```


#RFS by ctDNA MRD Window (2-16 weeks post-op)
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.MRD!="",]

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.MRD, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.MRD, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="RFS - ctDNA MRD Window", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", legend.labs=c("ctDNA Negative", "ctDNA Positive"), legend.title="")
circ_data$ctDNA.MRD <- factor(circ_data$ctDNA.MRD, levels=c("NEGATIVE","POSITIVE"))
summary(KM_curve, times= c(0, 24))
cox_fit <- coxph(surv_object ~ ctDNA.MRD, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)
```


#RFS by ctDNA post MRD
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.Surveillance!="",]

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.Surveillance, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.Surveillance, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="RFS - ctDNA Surveillance (post-MRD)", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", legend.labs=c("ctDNA Negative", "ctDNA Positive"), legend.title="")
circ_data$ctDNA.Surveillance <- factor(circ_data$ctDNA.Surveillance, levels=c("NEGATIVE","POSITIVE"))
summary(KM_curve, times= c(0, 24))
cox_fit <- coxph(surv_object ~ ctDNA.Surveillance, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)
```


#Time-dependent analysis for post-MRD (>16 weeks) time points
```{r}
rm(list=ls())
setwd("~/Downloads")
dt <- read_xlsx("NCC ESCC Time Dependent Data_postMRD.xlsx") |>
  clean_names() |>
  mutate(across(.cols = c(window_start_date,dfs_date,
                          surveillance_1_date:surveillance_12_date), 
                .fns = ~ as_date(as.Date(.x, format = "%Y-%m-%d"))))

dt_biomarker <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date,
         surveillance_1_status:surveillance_12_date) |>
  filter(ct_dna_surveillance_available) |>
  pivot_longer(cols = surveillance_1_status:surveillance_12_date,
               names_to = c("visit_number", ".value"),
               names_pattern = "surveillance_(.)_(.*)") |>
  mutate(biomarker_time = day(days(date - window_start_date))) |>
  select(pts_id, biomarker_time, biomarker_status = status) |>
  filter(!is.na(biomarker_time))

glimpse(dt_biomarker)

dt_survival <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date:dfs_date, dfs_event) |>  # Added dfs_event here
  filter(ct_dna_surveillance_available) |>
  mutate(dfs_time = (dfs_date - window_start_date),
         dfs_time = day(days(dfs_time)),
         dfs_event = as.numeric(dfs_event)) |>
  select(pts_id, dfs_time, dfs_event)

glimpse(dt_survival)

aux <- dt_survival %>% 
  filter(dfs_time <= 0)

tab <- left_join(aux, dt) |>
  select(pts_id, window_start_date, dfs_time, dfs_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = dfs_date:surveillance_12_date, 
                .fns = ~ as_date(.x))) |>
  select(pts_id, window_start_date, dfs_date, dfs_time)

datatable(tab, filter = "top")

dt_survival <- dt_survival |>
  filter(dfs_time > 0)

aux <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date, dfs_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = surveillance_1_date:surveillance_12_date, 
                .fns = ~ .x - window_start_date)) |>
  mutate(across(.cols = surveillance_1_date:surveillance_12_date, 
                .fns = ~ .x < 0)) |>
  rowwise() |>
  mutate(sum_neg = 
           sum(c_across(surveillance_1_date:surveillance_12_date),
               na.rm = TRUE))  |>
  select(pts_id, sum_neg)

tab <- left_join(aux, dt) |>
  filter(sum_neg > 0) |>
  select(pts_id, sum_neg, window_start_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = window_start_date:surveillance_12_date, 
                .fns = ~ as_date(.x))) 

datatable(tab, filter = "top")

aux <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date, dfs_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = dfs_date:surveillance_12_date, 
                .fns = ~ .x - window_start_date)) |>
  mutate(across(.cols = surveillance_2_date:surveillance_12_date,
                .fns = ~ dfs_date < .x)) |>
  rowwise() |>
  mutate(n_biomarker_after_event = sum(c_across(surveillance_2_date:
                                                  surveillance_12_date), 
                                       na.rm = TRUE)) |>
  mutate(across(.cols = surveillance_1_date:surveillance_12_date,
                .fns = ~ !is.na(.x))) |>
  mutate(total_biomarker = sum(c_across(surveillance_2_date:
                                          surveillance_12_date), 
                               na.rm = TRUE)) |>
  select(pts_id, n_biomarker_after_event, total_biomarker)

temp <- aux |> 
  select(-pts_id) |>
  group_by(n_biomarker_after_event, total_biomarker) |>
  summarize(freq = n())


tab <- left_join(aux, dt) |>
  select(pts_id, n_biomarker_after_event, total_biomarker, 
         dfs_date,
         surveillance_2_date:surveillance_12_date) |>
  mutate(across(.cols = dfs_date:surveillance_12_date, 
                .fns = ~ as_date(.x))) |>
  filter(n_biomarker_after_event > 0)
datatable(tab, filter = "top")

aux <- tmerge(data1 = dt_survival, 
              data2 = dt_survival,
              id = pts_id, 
              dfs_event = event(dfs_time, dfs_event))
dt_final <- tmerge(data1 = aux, 
                   data2 = dt_biomarker,
                   id = pts_id, 
                   biomarker_status = 
                     tdc(biomarker_time, biomarker_status))

datatable(dt_final, filter = "top")

# Syntax if there is not time-dependent covariate
# fit <- coxph(Surv(dfs_time, dfs_event) ~ biomarker_status,
#              data = dt_final)
# summary(fit)

fit <- coxph(Surv(tstart, tstop, dfs_event) ~ biomarker_status,
             data = dt_final)
summary(fit)
```


#RFS by ctDNA anytime post-surgery
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.anytime!="",]

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.anytime, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.anytime, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="RFS - ctDNA anytime post-surgery", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", legend.labs=c("ctDNA Negative", "ctDNA Positive"), legend.title="")
circ_data$ctDNA.anytime <- factor(circ_data$ctDNA.anytime, levels=c("NEGATIVE","POSITIVE"))
summary(KM_curve, times= c(0, 24))
cox_fit <- coxph(surv_object ~ ctDNA.anytime, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)
```


#Time-dependent analysis for post-surgery time points
```{r}
rm(list=ls())
setwd("~/Downloads")
dt <- read_xlsx("NCC ESCC Time Dependent Data_anytime postop.xlsx") |>
  clean_names() |>
  mutate(across(.cols = c(window_start_date,dfs_date,
                          surveillance_1_date:surveillance_12_date), 
                .fns = ~ as_date(as.Date(.x, format = "%Y-%m-%d"))))

dt_biomarker <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date,
         surveillance_1_status:surveillance_12_date) |>
  filter(ct_dna_surveillance_available) |>
  pivot_longer(cols = surveillance_1_status:surveillance_12_date,
               names_to = c("visit_number", ".value"),
               names_pattern = "surveillance_(.)_(.*)") |>
  mutate(biomarker_time = day(days(date - window_start_date))) |>
  select(pts_id, biomarker_time, biomarker_status = status) |>
  filter(!is.na(biomarker_time))

glimpse(dt_biomarker)

dt_survival <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date:dfs_date, dfs_event) |>  # Added dfs_event here
  filter(ct_dna_surveillance_available) |>
  mutate(dfs_time = (dfs_date - window_start_date),
         dfs_time = day(days(dfs_time)),
         dfs_event = as.numeric(dfs_event)) |>
  select(pts_id, dfs_time, dfs_event)

glimpse(dt_survival)

aux <- dt_survival %>% 
  filter(dfs_time <= 0)

tab <- left_join(aux, dt) |>
  select(pts_id, window_start_date, dfs_time, dfs_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = dfs_date:surveillance_12_date, 
                .fns = ~ as_date(.x))) |>
  select(pts_id, window_start_date, dfs_date, dfs_time)

datatable(tab, filter = "top")

dt_survival <- dt_survival |>
  filter(dfs_time > 0)

aux <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date, dfs_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = surveillance_1_date:surveillance_12_date, 
                .fns = ~ .x - window_start_date)) |>
  mutate(across(.cols = surveillance_1_date:surveillance_12_date, 
                .fns = ~ .x < 0)) |>
  rowwise() |>
  mutate(sum_neg = 
           sum(c_across(surveillance_1_date:surveillance_12_date),
               na.rm = TRUE))  |>
  select(pts_id, sum_neg)

tab <- left_join(aux, dt) |>
  filter(sum_neg > 0) |>
  select(pts_id, sum_neg, window_start_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = window_start_date:surveillance_12_date, 
                .fns = ~ as_date(.x))) 

datatable(tab, filter = "top")

aux <- dt |>
  select(pts_id, ct_dna_surveillance_available,
         window_start_date, dfs_date,
         surveillance_1_date:surveillance_12_date) |>
  mutate(across(.cols = dfs_date:surveillance_12_date, 
                .fns = ~ .x - window_start_date)) |>
  mutate(across(.cols = surveillance_2_date:surveillance_12_date,
                .fns = ~ dfs_date < .x)) |>
  rowwise() |>
  mutate(n_biomarker_after_event = sum(c_across(surveillance_2_date:
                                                  surveillance_12_date), 
                                       na.rm = TRUE)) |>
  mutate(across(.cols = surveillance_1_date:surveillance_12_date,
                .fns = ~ !is.na(.x))) |>
  mutate(total_biomarker = sum(c_across(surveillance_2_date:
                                          surveillance_12_date), 
                               na.rm = TRUE)) |>
  select(pts_id, n_biomarker_after_event, total_biomarker)

temp <- aux |> 
  select(-pts_id) |>
  group_by(n_biomarker_after_event, total_biomarker) |>
  summarize(freq = n())


tab <- left_join(aux, dt) |>
  select(pts_id, n_biomarker_after_event, total_biomarker, 
         dfs_date,
         surveillance_2_date:surveillance_12_date) |>
  mutate(across(.cols = dfs_date:surveillance_12_date, 
                .fns = ~ as_date(.x))) |>
  filter(n_biomarker_after_event > 0)
datatable(tab, filter = "top")

aux <- tmerge(data1 = dt_survival, 
              data2 = dt_survival,
              id = pts_id, 
              dfs_event = event(dfs_time, dfs_event))
dt_final <- tmerge(data1 = aux, 
                   data2 = dt_biomarker,
                   id = pts_id, 
                   biomarker_status = 
                     tdc(biomarker_time, biomarker_status))

datatable(dt_final, filter = "top")

# Syntax if there is not time-dependent covariate
# fit <- coxph(Surv(dfs_time, dfs_event) ~ biomarker_status,
#              data = dt_final)
# summary(fit)

fit <- coxph(Surv(tstart, tstop, dfs_event) ~ biomarker_status,
             data = dt_final)
summary(fit)
```


#RFS by ctDNA Dynamics post-NAC to MRD - 4 groups
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")

circ_data$ctDNA.Dynamics <- NA #first we create the variable for the ctDNA & NAC combination, and we assign values
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.postnac == "NEGATIVE" & ctDNA.MRD == "NEGATIVE" ~ "1",
    ctDNA.postnac == "POSITIVE" & ctDNA.MRD == "NEGATIVE" ~ "2",
    ctDNA.postnac == "NEGATIVE" & ctDNA.MRD == "POSITIVE" ~ "3",
    ctDNA.postnac == "POSITIVE" & ctDNA.MRD == "POSITIVE" ~ "4"
  ))

circ_data <- circ_data[!is.na(circ_data$ctDNA.Dynamics),]
survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.Dynamics, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.Dynamics, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","green","purple","red"), title="RFS - ctDNA Dynamics post-NAC - MRD Window", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", legend.labs=c("Persistently Negative","Converted Negative", "Converted Positive","Persistently Positive"), legend.title="")
summary(KM_curve, times= c(0, 24))
circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2","3","4"), labels=c("Persistently Negative","Converted Negative","Converted Positive","Persistently Positive"))
cox_fit <- coxphf(surv_object ~ ctDNA.Dynamics, data=circ_data)
summary(cox_fit)
```


#Association of ctDNA post-NAC and TRG status
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")

#Vertical Fisher plot for ctDNA clearance post-NAT and Rec Status
circ_data <- circ_data[!is.na(circ_data$ctDNA.postnac),]
circ_data$ctDNA.postnac <- factor(circ_data$ctDNA.postnac, levels=c("NEGATIVE","POSITIVE"))
circ_data$TRG <- factor(circ_data$TRG, levels = c("1", "2", "3"), labels = c("TRG1", "TRG2", "TRG3"))
contingency_table <- table(circ_data$TRG, circ_data$ctDNA.postnac)
fisher_exact_test <- fisher.test(contingency_table)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2

# Swapping x and y in ggplot function to make bar plot vertical
ggplot(table_df, aes(y = Var1, x = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(x = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA post-NAT & TRG", y = "TRG Status", x = "Patients (%)", fill = "ctDNA post-NAC") +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("NEGATIVE" = "lightblue3", "POSITIVE" = "red")) + # define custom colors
  theme(axis.text.y = element_text(angle = 0, hjust = 1.5, size = 14), # increase y-axis text size
        axis.text.x = element_text(size = 14, color = "black"), # increase x-axis text size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Recurrence label size
```


#RFS by ctDNA post-NAC and TRG - 3 groups
```{r}
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.postnac!="",]

circ_data$ctDNA.pCR <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.pCR = case_when(
    ctDNA.postnac == "NEGATIVE" & TRG %in% c(1, 2) ~ "1",
    ctDNA.postnac == "NEGATIVE" & TRG == 3 ~ "2",
    ctDNA.postnac == "POSITIVE" & TRG %in% c(1, 2) ~ "3"
  ))

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.pCR, data = circ_data)
circ_data <- circ_data[!is.na(circ_data$ctDNA.pCR),]
circ_data$ctDNA.pCR <- factor(circ_data$ctDNA.pCR, levels=c("1","2", "3"), labels=c("TRG1/2 ctDNA (-)","TRG3 ctDNA (-)", "TRG1/2 ctDNA (+)"))
survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.pCR, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.pCR, data = circ_data,conf.int=0.95,conf.type="log-log")
summary(KM_curve, times= c(0, 24))
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, 
           break.time.by=6, palette=c("blue", "green", "red"), title="DFS - ctDNA post-NAC & TRG", ylab= "Recurrence-Free Survival", xlab="Months from Surgery", 
           legend.labs=c("TRG1/2 ctDNA (-)","TRG3 ctDNA (-)", "TRG1/2 ctDNA (+)"), legend.title="")

rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.postnac!="",]

circ_data$ctDNA.pCR <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.pCR = case_when(
    ctDNA.postnac == "NEGATIVE" & TRG %in% c(1, 2) ~ "1",
    ctDNA.postnac == "NEGATIVE" & TRG == 3 ~ "2",
    ctDNA.postnac == "POSITIVE" & TRG %in% c(1, 2) ~ "3"
  ))
circ_data$ctDNA.pCR <- factor(circ_data$ctDNA.pCR, levels=c("1","2","3"))
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
cox_fit <- coxph(surv_object ~ ctDNA.pCR, data=circ_data)
summary(cox_fit)
```


#RFS by ctDNA post-NAC and pCR - 3 groups
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.4w!="",]
circ_data$RFS.months=circ_data$RFS.months-1
circ_data <- circ_data[circ_data$RFS.months>=0,]

circ_data$ctDNA.pCR <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.pCR = case_when(
    ctDNA.postnac == "NEGATIVE" & pCR == "TRUE" ~ "1",
    ctDNA.postnac == "NEGATIVE" & pCR == "FALSE" ~ "2",
    ctDNA.postnac == "POSITIVE" & pCR == "FALSE" ~ "3",
  ))

survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.pCR, data = circ_data)
circ_data <- circ_data[!is.na(circ_data$ctDNA.pCR),]
circ_data$ctDNA.pCR <- factor(circ_data$ctDNA.pCR, levels=c("1","2", "3"), labels=c("pCR ctDNA (-)","No pCR ctDNA (-)", "No pCR ctDNA (+)"))
survfit(Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)~ctDNA.pCR, data = circ_data)
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.pCR, data = circ_data,conf.int=0.95,conf.type="log-log")
summary(KM_curve, times= c(0, 24))
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, 
           break.time.by=6, palette=c("blue", "green", "red"), title="DFS - ctDNA post-NAC & pCR", ylab= "Recurrence-Free Survival", xlab="Time (Months)", 
           legend.labs=c("pCR ctDNA (-)","No pCR ctDNA (-)", "No pCR ctDNA (+)"), legend.title="")

rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.4w!="",]
circ_data$RFS.months=circ_data$RFS.months-1
circ_data <- circ_data[circ_data$RFS.months>=0,]

circ_data$ctDNA.pCR <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.pCR = case_when(
    ctDNA.postnac == "NEGATIVE" & pCR == "TRUE" ~ "1",
    ctDNA.postnac == "NEGATIVE" & pCR == "FALSE" ~ "2",
    ctDNA.postnac == "POSITIVE" & pCR == "FALSE" ~ "3",
  ))
circ_data$ctDNA.pCR <- factor(circ_data$ctDNA.pCR, levels=c("1","2","3"))
surv_object <-Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event)
cox_fit <- coxph(surv_object ~ ctDNA.pCR, data=circ_data)
summary(cox_fit)
```


#Multivariate cox regression for RFS - ctDNA post-NAC
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.postnac!="",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.postnac <- factor(circ_data$ctDNA.postnac, levels=c("NEGATIVE","POSITIVE"), labels = c("Negative", "Positive"))
circ_data$Age.Group <- factor(circ_data$Age.Group, levels=c("1","2"), labels = c("≤70", ">70"))
circ_data$Stage.Group <- factor(circ_data$Stage.Group, levels = c("I-II", "III-IV"))
surv_object <- Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event) 
cox_fit <- coxph(surv_object ~ ctDNA.postnac + Age.Group + Stage.Group, data=circ_data) 
ggforest(cox_fit, data = circ_data, main = "Multivariate Regression Model for RFS", refLabel = "Reference Group")
test.ph <- cox.zph(cox_fit)
```


#Multivariate cox regression for RFS - ctDNA MRD Window
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("NCC ESCC Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.MRD!="",]

circ_data$ctDNA.MRD <- factor(circ_data$ctDNA.MRD, levels=c("NEGATIVE","POSITIVE"), labels = c("Negative", "Positive"))
circ_data$Age.Group <- factor(circ_data$Age.Group, levels=c("1","2"), labels = c("≤70", ">70"))
circ_data$Stage.Group <- factor(circ_data$Stage.Group, levels = c("I-II", "III-IV"))
surv_object <- Surv(time = circ_data$RFS.months, event = circ_data$RFS.Event) 
cox_fit <- coxph(surv_object ~ ctDNA.MRD + Age.Group + Stage.Group, data=circ_data) 
ggforest(cox_fit, data = circ_data, main = "Multivariate Regression Model for RFS", refLabel = "Reference Group")
test.ph <- cox.zph(cox_fit)
```

